---
title: "Technical Labs"
format: html
editor: visual
---

# Landscape Connectivity Analysis for Species Movement

This project applied landscape ecology principles to evaluate structural and functional connectivity across a heterogeneous landscape using spatial analysis and network-based modeling in R. Using National Land Cover Data, landscape configuration was analyzed to understand how habitat distribution and matrix composition influence species movement and connectivity. Landscape metrics were used to quantify patch structure and spatial patterns, while network analysis techniques were employed to model potential movement pathways across the landscape. The project incorporated resistance-based modeling to simulate species-specific movement scenarios for ducks and spotted owls, accounting for differences in habitat preference and dispersal ability. Graphs and grains of connectivity models were developed across multiple thresholds to assess connectivity sensitivity at different spatial scales. The results highlight how landscape structure, habitat quality, and movement assumptions interact to shape ecological connectivity, demonstrating the value of connectivity modeling for conservation planning and landscape-scale environmental decision-making

**Key methods and Data**

Landscape ecology and connectivity theory, Structural and functional connectivity analysis, Network analysis, National Land Cover Data (NLCD)

**Tools**

Rstudio (landscapemetrics- Core area metric, Aggregation metric, Area and Edge metric)

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=14, fig.height=6}
library(png)

# Absolute file paths for P1 and P2
image_paths <- c(
  "C:/Users/kb90.stu/OneDrive - UBC/Documents/GitHub/E-Portfolio/images/p1.png",
  "C:/Users/kb90.stu/OneDrive - UBC/Documents/GitHub/E-Portfolio/images/p2.png"
)

# Corresponding captions
captions <- c(
  "Least Cost Connectivity of Owl Habitat Patches",
  "Landscape Resistance Map for Owl Habitat"
)

# Read images
imgs <- lapply(image_paths, readPNG)

# Layout: 1 row x 2 columns
par(mfrow = c(1,2), mar = c(3,2,3,2))  # margins for captions

# Loop through images and plot
for(i in 1:2){
  plot(0:1, 0:1, type="n", axes=FALSE, ann=FALSE)   # empty plot
  rasterImage(imgs[[i]], 0, 0, 1, 1)               # fill the cell
  mtext(bquote(bold(.(captions[i]))), side=3, line=1, cex=1.1)  # bold + larger  # caption above image
}

```

# Analysis of Fire Burn Severity and Post-Fire Vegetation Recovery Using Landsat Time-Series

The Prouton Lakes wildfire (2017) in the Alex Fraser Research Forest, BC, was examined for burn intensity and post-fire vegetation recovery using Landsat 8 surface reflectance time-series (2013-2021). Burn severity was assessed using the normalized burn ratio (NBR) and delta NBR (dNBR), which allowed impacted areas to be classified as unburned, low, medium, or high severity. Vegetation recovery was assessed by observing NDVI trends across burn severity classes for four years after the fire. Time-series composites, supervised classifiers, and statistical visualizations were used to investigate regional patterns of fire effects and vegetation recovery. This study highlights the use of remote sensing and spatial analysis in landscape ecology to aid in wildfire impact assessment and ecosystem management decisions.

**Key methods and Data**

Remote sensing analysis using Landsat 8 OLI Level-2 surface reflectance data, Vegetation and burn indices: NDVI, NBR, dNBR, Time-series processing and compositing for multiple years, Raster pre-processing: masking, cropping, and QA filtering

**Tools**

Rstudio (raster/terra, ggplot2)

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=14, fig.height=6}
library(png)

# Absolute file paths for P1 and P2
image_paths <- c(
  "C:/Users/kb90.stu/OneDrive - UBC/Documents/GitHub/E-Portfolio/images/rs3.png",
  "C:/Users/kb90.stu/OneDrive - UBC/Documents/GitHub/E-Portfolio/images/rs4.png",
  "C:/Users/kb90.stu/OneDrive - UBC/Documents/GitHub/E-Portfolio/images/rs2.png",
  "C:/Users/kb90.stu/OneDrive - UBC/Documents/GitHub/E-Portfolio/images/rs1.png"
)

# Corresponding captions
captions <- c(
  "Yearly Mean Normalized Burn Ratio (NBR)",
  "Yearly Mean Normalized Diffrential Vegetation Index (NDVI)",
  "Prouton Lakes Fire Burn Severity",
  "NDVI Recovery at Different Burn Severity Levels"
)

# Read images
imgs <- lapply(image_paths, readPNG)

# Layout: 1 row x 2 columns
par(mfrow = c(1,2), mar = c(3,2,3,2))  # margins for captions

# Loop through images and plot
for(i in 1:4){
  plot(0:1, 0:1, type="n", axes=FALSE, ann=FALSE)   # empty plot
  rasterImage(imgs[[i]], 0, 0, 1, 1)               # fill the cell
  mtext(bquote(bold(.(captions[i]))), side=3, line=1, cex=1.1)  # bold + larger  # caption above image
}

```

# Cartographic Modelling of Old-Growth Forests on Vancouver Island With Forest Inventories

Although old-growth forests provide important ecological, biological, and wood resources, they are under threat from logging and human disturbances. This lab investigates the distribution of prospective old-growth stands on Vancouver Island, British Columbia, utilizing VRI data, land ownership layers, human disturbance data, and landscape units. Forest polygons were filtered to include managed crown forests, which were then divided into Seral Stages based on age and BEC zone, and human disturbances were adjusted. The percentage of old-growth forest was estimated for each Landscape Unit-BEC combination and compared to provincial targets. Cartographic modelling was used to depict the data, resulting in maps and tables that show locations that exceeded or fell short of the old-growth target. This workflow demonstrates how GIS and Python-based spatial analysis can help with forest management and conservation planning.

**Key methods and Data**

Subsetting VRI polygons to managed forests and extracting relevant attributes, Identifying crown forests using land ownership and human disturbance layers, Assigning Seral Stages (Early, Mid, Mature, Old) based on BEC zones and projected stand age; resetting stages in disturbed areas, Vegetation Resource Inventory 2024 (vancouver_island_vri), Generalized Forest Cover Ownership (vancouver_island_own)

**Tools**

ArcGIS Pro, Python, SQL

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=10, fig.height=8}
library(png)

# File path for the image
image_path <- "C:/Users/kb90.stu/OneDrive - UBC/Documents/GitHub/E-Portfolio/images/gis1.png"

# Caption
caption <- "Forested Areas of Vancouver Island Showing Threshold of Biogeoclimatic (BEC) Subzones"

# Read image
img <- readPNG(image_path)

# Layout: 1 row x 1 column
par(mfrow = c(1,1), mar = c(5,2,5,2))  # extra margin for bold caption

# Plot image
plot(0:1, 0:1, type="n", axes=FALSE, ann=FALSE)
rasterImage(img, 0, 0, 1, 1)

# Add bold caption above the image
mtext(bquote(bold(.(caption))), side=3, line=1, cex=1.2)

```

# Machine Learning Techniques to Classify Deciduous or Coniferous Trees in the Petawawa Research Forest

Accurate mapping of coniferous and deciduous tree species is essential for effective forest management, ecosystem monitoring, and long-term environmental planning. These forest types differ in growth patterns, ecological functions, and carbon storage potential, making their spatial delineation critical for decision-making. Advances in remote sensing and machine learning have created new opportunities to classify forest composition using spectral and structural information derived from satellite and aerial imagery. This project evaluated the performances of machine-learning classifiers (Random Forest (RF) and Support Vector Machine (SVM)), for distinguishing coniferous and deciduous trees within the Petawawa Research Forest. Both models were trained using the same set of predictor variables and reference data so that their performance could be compared fairly. Accuracy was assessed using classification metrics, including overall accuracy, precision, recall, and confusion matrices, along with model computation time. This direct comparison highlights how Random Forest and Support Vector Machine differ in both classification performance and efficiency, providing practical insight into their suitability for conifer and deciduous tree mapping in Canadian forest landscapes

**Key methods and Data**

Supervised machine learning classification of forest cover into coniferous and deciduous classes, Training and testing of two classifiers: Random Forest and Support Vector Machine, Comparison of model performance based on both classification accuracy and computation time, Airborne LiDAR 

**Tools**

Arcgis Pro, Python, Scikit learn

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=10, fig.height=8}
library(png)

# File path for the image
image_path <- "C:/Users/kb90.stu/OneDrive - UBC/Documents/GitHub/E-Portfolio/images/Py1.png"

# Caption
caption <- "Comparing SVM and RF classifiers performance on coniferous and deciduous tree species"

# Read image
img <- readPNG(image_path)

# Layout: 1 row x 1 column
par(mfrow = c(1,1), mar = c(5,2,5,2))  # extra margin for bold caption

# Plot image
plot(0:1, 0:1, type="n", axes=FALSE, ann=FALSE)
rasterImage(img, 0, 0, 1, 1)

# Add bold caption above the image
mtext(bquote(bold(.(caption))), side=3, line=1, cex=1.2)

```

