---
title: " Capstone Project"
format: html
editor: visual
---

## Unmanned Aerial Vehicle (UAV) Multispectral Imagery and GIS-based Approaches for Vineyard Row Identification and Segmentation in Precision Viticulture

Spatial variability within vineyards significantly influences grapevine growth, yield, and fruit quality, making accurate monitoring of vine vigour essential for effective management. Advances in precision viticulture have enabled the integration of remote sensing and geospatial technologies to analyze vineyard conditions at fine spatial scales. This capstone project applies UAV-based multispectral imagery and geospatial analysis to improve the identification and segmentation of vineyard rows in a heterogeneous vineyard environment in the Okanagan region. Vegetation indices, including the Normalized Difference Red Edge (NDRE) and Green Normalized Difference Vegetation Index (GNDVI), were derived to assess canopy vigour and spatial variability across the vineyard. Image analysis techniques were used to separate vine canopy from inter-row vegetation, soil, and shadow to ensure accurate extraction of canopy statistics. Zonal statistical analysis was then performed on the segmented vine areas to quantify spatial patterns of vine health and growth uniformity. The results demonstrate how improved canopy delineation enhances the reliability of vegetation index analysis, providing more accurate spatial information to support data-driven decision-making in precision vineyard management.

**Key methods and Data**

Canopy Height Model, Vines Thresholding, Segmentation of vine rows, Zonal Statistics and NDRE of vines with GNDVI comparison, UAV Multispectral Data

**Tools**

QGIS, Rstudio

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=14, fig.height=6}
library(jpeg)

# Absolute file paths for P1 and P2
image_paths <- c(
  "C:/Users/kb90.stu/OneDrive - UBC/Documents/GitHub/E-Portfolio/images/seg.jpeg",
  "C:/Users/kb90.stu/OneDrive - UBC/Documents/GitHub/E-Portfolio/images/ndre_zonal.jpeg",
  "C:/Users/kb90.stu/OneDrive - UBC/Documents/GitHub/E-Portfolio/images/gndvi_zonal.jpeg",
  "C:/Users/kb90.stu/OneDrive - UBC/Documents/GitHub/E-Portfolio/images/area.jpeg"
)

# Corresponding captions
captions <- c(
  "Segmentated vines of both test vine block (A) and applied vine block (B)",
  "NDRE zonal statistics map on vine blocks (A/B)",
  "GNDVI zonal statisitcs map on vine blocks (A/B)",
  "Total vine canopy area per row segment on vine blocks (A/B)"
)

# Read images
imgs <- lapply(image_paths, readJPEG)

# Layout: 1 row x 2 columns
par(mfrow = c(1,2), mar = c(3,2,3,2))  # margins for captions

# Loop through images and plot
for(i in 1:4){
  plot(0:1, 0:1, type="n", axes=FALSE, ann=FALSE)   # empty plot
  rasterImage(imgs[[i]], 0, 0, 1, 1)               # fill the cell
  mtext(bquote(bold(.(captions[i]))), side=3, line=1, cex=1.1)  # bold + larger  # caption above image
}

```

## Crime Analysis and Assessment using Data Mining Techniques

Crimes are actions committed in violation of the law. These crime activities have increased over the years due to high standards of living, unemployment, etc. which need to be curbed to ensure a safe society. Law enforcement agencies can effectively curb the canker by analysing previous crimes to understand the pattern of occurrence of these crimes to gather intelligence to tackle the situation effectively. Initially, law enforcement agencies carry out analysis of crimes on papers or use analogue techniques which is time consuming and tedious as one may have to deal with large paper sizes. In recent years, digital techniques such as the use of artificial intelligence which have proven to be better than traditional techniques are been used to carry out various analyses of crime data. This study has employed digital techniques such as data mining to analyse crime data in Chicago. The techniques used were the Random Forest classifier and Decision tree which were used to analyse and classify crimes into low, medium and high crimes. In comparison to the Decision tree, the Random Forest classifier outperformed it. Accuracy, precision, and recall for the Random Forest classifier are, respectively, 77%, 77%, and 77%. However, the decision trees' values for accuracy, precision, and recall are 75%, 75%, and 76%, respectively. Additionally, a map of crime hotspots was created using an analysis of Chicago's historical crime data to show security agencies where crimes are most likely to occur

**Key methods and Data**

Data mining techniques (decision trees and random forest classifiers) for crime prediction and analysis, Spatial clustering, Secondary data obtained from Chicago Police Department Statistics Portal

**Tools**

Openstreetmap, Google colab

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=14, fig.height=6}
library(png)

# Absolute file paths for P1 and P2
image_paths <- c(
  "C:/Users/kb90.stu/OneDrive - UBC/Documents/GitHub/E-Portfolio/images/C1.png",
  "C:/Users/kb90.stu/OneDrive - UBC/Documents/GitHub/E-Portfolio/images/C2.png",
  "C:/Users/kb90.stu/OneDrive - UBC/Documents/GitHub/E-Portfolio/images/C3.png",
  "C:/Users/kb90.stu/OneDrive - UBC/Documents/GitHub/E-Portfolio/images/C4.png"
)

# Corresponding captions
captions <- c(
  " Crimes Location on a Map",
  "Hotspot Areas of Crimes",
  "Crimes Heatmap",
  "Comparison of the Two Classification Techniques"
)

# Read images
imgs <- lapply(image_paths, readPNG)

# Layout: 1 row x 2 columns
par(mfrow = c(1,2), mar = c(3,2,3,2))  # margins for captions

# Loop through images and plot
for(i in 1:4){
  plot(0:1, 0:1, type="n", axes=FALSE, ann=FALSE)   # empty plot
  rasterImage(imgs[[i]], 0, 0, 1, 1)               # fill the cell
  mtext(bquote(bold(.(captions[i]))), side=3, line=1, cex=1.1)  # bold + larger  # caption above image
}

```
